---
title: "Retrieving Taxonomic Info"
author: "Courtney Stepien"
date: "August 4, 2016"
output: 
 html_document:
    toc: true
    number_sections: true
    theme: cerulean
    highlight: pygments
    fig_caption: yes
    self_contained: false
---

# File goals

We will use the packages taxize and myTAI to get higher-level taxonomy for my genera, so we can table basic stats about the number of taxa per family, order and class. 

We will use a number of different taxonomy databases: 

* Barcode of Life Data Systems (BOLD) <- taxize package
* Encyclopedia of Life (EOL) <- taxize package
* Integrated Taxonomic Information Service (ITIS) <- taxize package
* National Center for Biotechnology Information (NCBI) <- taxize package
* World Register of Marine Species (WoRMS) <- taxizesoap package

[taxize package github link](https://github.com/ropensci/taxize) for tutorials

[taxizesoap package github link](https://github.com/ropensci/taxizesoap) for tutorials

# Next steps

Next steps are:

1. load the dataset
2. get a genus list
3. figure out how to use taxise and taxisesoap to get higher-level taxonomy info 

# Analysis

## Data setup and packages

```{r, message=FALSE, warning=FALSE}
library(taxize)
library(taxizesoap)
library(dplyr)
#data <- read.csv("../data/input_filename.csv")
```

```{r, echo=FALSE}

```
